Supporting Information for\\ ``A weakly structured stem for human origins in Africa''

### 2022-03-11 supplement.tex
more modern to present changes
formatting

\usepackage[margin=1in]{geometry}
\usepackage{listings}
\usepackage{url}
\usepackage{authblk}
\usepackage{graphicx}
\usepackage{color}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{xspace}

@@

\begin{document}
\title{Supporting Information for\\
``A weakly structured stem for human origins in Africa''}
\author{Aaron P. Ragsdale, Timothy D. Weaver, Elizabeth G. Atkinson, Eileen Hoal, Marlo M\"{o}ller, Brenna M. Henn, and Simon Gravel}
\date{\today}
\maketitle

@@

Here we focus on four parameterizations of early human history: (1) a “single
origin” with recent expansion, (2) primary and second stem populations with
“continuous migration”, (3) two stem populations that do not exchange migrants
before merging to form regional African populations, referred to as a “merger
without stem migration”, and (4) two stem populations that are allowed to
exchange migrants before merging to form regional African populations, referred
to as a “merger with stem migration”. Given the best fit parameters for each
model
(Tables~\ref{tab:supp-single-origin}--\ref{tab:supp-merger-with-stem-migration}),
we used \msprime version 1.0.4 \citep{Kelleher2016-lw,Baumdicker2021-mj} to
simulate sampled individual genomes to match the data used in inference (see
simulation details below). We used these simulations to compare our model predictions to 
predictions based on the conditional SFS (cSFS) and on relate curves (see below).

@@

Recent studies have focused on characterizing the timing of the earliest
divergences among human populations, and such analyses
typically fit simple isolation-with-migration (IM) or ``clean split'' models to
observed allele frequencies or inferred coalescence times among and between
populations (see \citet{Weaver2008-ho,Bergstrom2021-iw} for recent reviews).
Simple IM models
typically fit a constant $N_e$ within the two diverged population, the split
time, and a symmetric continuous migration rate (which could be fixed to zero,
equivalent to the ``clean split'' model). We explored the effect of fitting
such simplified models of history to data from more complex models by
simulating the site frequency spectrum (SFS) under each of our four models for
pairs of populations and reinferring an IM and clean split model for each pair.
The simulations sampled 10 diploid individuals from each
population, and we used \msprime \citep{Baumdicker2021-mj} to simulated
500 Mb of total sequence length (the sum over500 replicates each 1 Mb in length)
with constant recombination rate $r=10^{-8}$ and mutation rate $u=2\times10^{-8}.$

>>>>

\usepackage[margin=1in]{geometry}
\usepackage{listings}
\usepackage{url}

\usepackage{authblk}
\renewcommand{\Affilfont}{\normalsize}
\renewcommand{\Authfont}{\normalsize}

\usepackage{graphicx}
\usepackage{color}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{xspace}

@@

\begin{document}
\title{Supporting Information for\\
``A weakly structured stem for human origins in Africa''}
\author{Aaron P. Ragsdale, Timothy D. Weaver, Elizabeth G. Atkinson, Eileen Hoal, Marlo M\"{o}ller, Brenna M. Henn, and Simon Gravel}
\date{\normalsize \today}
\maketitle

@@

Here we focus on four parameterizations of early human history: (1) a “single
origin” with recent expansion, (2) primary and second stem populations with
“continuous migration”, (3) two stem populations that do not exchange migrants
before merging to form regional African populations, referred to as a “merger
without stem migration”, and (4) two stem populations that are allowed to
exchange migrants before merging to form regional African populations, referred
to as a “merger with stem migration”. Given the best fit parameters for each
model
(Tables~\ref{tab:supp-single-origin}--\ref{tab:supp-merger-with-stem-migration}),
we used \msprime version 1.0.4 \citep{Kelleher2016-lw,Baumdicker2022-mj} to
simulate sampled individual genomes to match the data used in inference (see
simulation details below). We used these simulations to compare our model predictions to 
predictions based on the conditional SFS (cSFS) and on relate curves (see below).

@@

Recent studies have focused on characterizing the timing of the earliest
divergences among human populations, and such analyses
typically fit simple isolation-with-migration (IM) or ``clean split'' models to
observed allele frequencies or inferred coalescence times among and between
populations (see \citet{Weaver2008-ho,Bergstrom2021-iw} for recent reviews).
Simple IM models
typically fit a constant $N_e$ within the two diverged population, the split
time, and a symmetric continuous migration rate (which could be fixed to zero,
equivalent to the ``clean split'' model). We explored the effect of fitting
such simplified models of history to data from more complex models by
simulating the site frequency spectrum (SFS) under each of our four models for
pairs of populations and reinferring an IM and clean split model for each pair.
The simulations sampled 10 diploid individuals from each
population, and we used \msprime \citep{Baumdicker2022-mj} to simulated
500 Mb of total sequence length (the sum over500 replicates each 1 Mb in length)
with constant recombination rate $r=10^{-8}$ and mutation rate $u=2\times10^{-8}.$
